# Reproducible Research Course Project 1
Sarah Guo
7/26/2017

## Loading and preprocessing the data
```{r setoptions, echo=TRUE}
library(knitr)
opts_chunk$set(echo=TRUE)
```
```{r loaddata}

setwd("D:/Google Drive/stats begins here/Data Science -John Hopkins/Course 5 Reproducible Research/Week 2 lecture/Project 1")
fileUrl<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="activity.zip")
dateDownloaded<-date()
unzip("activity.zip")

act<-read.csv("activity.csv",header=TRUE)
```
The dataset was downloaded on `r dateDownloaded`.

## What is mean total number of steps taken per day?

**Here is a histogram of the total number of steps taken each day.**

```{r}
eachday<-aggregate(steps~date,act,sum,na.rm=TRUE)
library(ggplot2)
qplot(steps,data=eachday,binwidth=500,xlab="Steps",ylab="Frequency")+labs(title="Total Number of Steps Taken Each Day")
```

**Mean and median of the total number of steps taken per day**

```{r calculatemean}
meansteps<-mean(eachday$steps,na.rm = TRUE)
mediansteps<-median(eachday$steps,na.rm = TRUE)
```
*The mean of the total number of steps taken per day was `r meansteps`.*

*The median of the total number of steps taken per day was `r mediansteps`.*

## What is the average daily activity pattern?

**Here is a time series plot showing average daily activity pattern.**

```{r}
avinterval<-aggregate(steps~interval,act,mean,na.rm=TRUE)
ggplot(avinterval,aes(interval,steps))+geom_line(col="blue")+xlab("5-minute interval")+ylab("average number of steps taken")+labs(title="Average Daily Activity Pattern")
```

**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**
```{r maxsteps}
minterval<- avinterval[which.max(avinterval$steps),]
kable(minterval,digits=1)
```

##Imputing missing values

**Calculate total number of missing values in the dataset**

```{r missingvalues}
N<-sum(is.na(act$steps))
```
*The total number of missing values in the dataset was `r N`.*

**Devise a strategy for filling in all of the missing values in the dataset.Create a new dataset that is equal to the original dataset but with the missing data filled in with column mean.**

```{r fillinmissing}
act1<-act
act1$steps[which(is.na(act1$steps))]<-mean(act1$steps,na.rm=TRUE)
```

**Here is a histogram of the total number of steps taken each day using the new dataset.**

```{r}
eachday1<-aggregate(steps~date,act1,sum)
qplot(steps,data=eachday1,binwidth=500,xlab="Steps",ylab="Frequency")+labs(title="Total Number of Steps Taken Each Day")
```

**mean and median of the total number of steps taken per day**
```{r}
meansteps1<-mean(eachday1$steps)
mediansteps1<-median(eachday1$steps)
```
*The mean of the total number of steps taken per day was `r meansteps1`*

*The median of the total number of steps taken per day was `r mediansteps1`*

*The mean and median are different from the estimates from the first part of the assignment.*

*The mean is equal to median after filling in the missing values.*

## Are there differences in activity patterns between weekdays and weekends?

**Create a new factor - weekend/weekday**

```{r newfactor}
library(dplyr)
act1$date<-as.Date(act1$date,"%Y-%m-%d")
act1<- act1%>%
  mutate(weektype= ifelse(weekdays(act1$date)=="Saturday" | weekdays(act1$date)=="Sunday", "Weekend", "Weekday"))
```

**Make a ggplot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)**

```{r}
avinterval1<-aggregate(steps~interval+weektype,act1,mean)
ggplot(avinterval1,aes(interval,steps,color=weektype))+geom_line()+facet_grid(weektype~.)+xlab("5-minute interval")+ylab("average number of steps")+labs(title="Daily Activity Patterns on Weekdays and Weekends")
```













